
```{r}
suppressPackageStartupMessages(library(ggplot2))
suppressMessages(library(dplyr))
```

```{r}
df <- read.csv("titanic_train.csv", stringsAsFactors = FALSE)
df$Age[is.na(df$Age)] <- median(df$Age, na.rm = TRUE)
df$Pclass <- ordered(df$Pclass, levels = c('1', '2', '3'))
df$Survived <- factor(df$Survived)
df$Embarked[df$Embarked == ''] <- 'S'   # Remove the empty char 
df <- select(df, "Survived", "Pclass", "Sex", "Age", "Fare", "Embarked")
```


# BarCharts

There are two types functions in ggplot2 for bar charts : 
- `geom_bar` makes the height of the bar proportional to the number of cases in each group (or if the weight aesthetic is supplied, the sum of the weights). It uses stat_count by default, it counts the number of cases at each x position

- `geom_col` makes the heights of the bars to represent values in the data,  It uses stat_identity, it leaves the data as is.

### Aesthetics

`geom_bar` understands the following aesthetics:

- x
- y
- alpha
- color
- fill
- group
- linetype
- size

## Bargraph to show the number of passengers embarked from each port

```{r}
ggplot(data = df, aes(x = Embarked)) +
  geom_bar()
```

## Bargraph to show the port which had the paid the maximum total fare paid by its corresponding passenegers

```{r}
ggplot(data = df, aes(x = Embarked)) +
  geom_bar(aes(weight = Fare))
```

## Stacked Bar Plot

Bar charts are automatically stacked when multiple bars are placed at the same location. The order of the fill is designed to match the legend.

```{r}
ggplot(data = df, aes(x = Embarked)) +
  geom_bar(aes(fill = Pclass))
```

To change the order in which the multiple stacks are placed on the bar use 
`position = position_stack(reverse = TRUE)`

```{r}
ggplot(data = df, aes(x = Embarked)) +
  geom_bar(aes(fill = Pclass), position = position_stack(reverse = TRUE)) +
  scale_fill_brewer(palette = "Pastel1") 
```

## Horizontal bargraph (Obtained by flipping the coordinates)

```{r}
ggplot(data = df, aes(x = Embarked)) +
  geom_bar(aes(fill = Pclass), position = position_stack(reverse = TRUE)) +
  coord_flip() +
  scale_fill_brewer(palette = "Pastel1") +
  theme_minimal()
```

To arrange the bar plots in ascending or descending order.

```{r}

```

## Grouped bar charts



```{r}
ggplot(data = df, aes(x = Embarked, fill = Survived)) +
  geom_bar(position = "dodge")
```

It is important to use `position = "dodge"` otherwise it would result in a 
stacked bar graph. `position = "dodge"`seperates the bar based on the __fill__ category.




