------------------------------------------------------
# Simpson's Paradox
------------------------------------------------------

```{r}
df <- data.frame(c("Day 1", "Day 1", "Day 1", "Day 1", "Day 1",
                   "Day 2", "Day 2", "Day 2", "Day 2", "Day 2"),
                 c("T1", "T2", "T3", "T4", "T5", 
                   "T1", "T2", "T3", "T4", "T5"),
                 c("Bob", "John", "Bob", "Bob", "Bob", 
                   "John", "John", "Bob", "John", "John"),
                 c("Pending", "Pending", "Resolved", "Pending", "Pending",
                   "Resolved", "Resolved", "Resolved", "Pending", "Resolved"));
names(df) <- c("Day", "Ticket No.", "Assignee", "Status")
df
```


 Overall Result
```{r}
table(df$Assignee, df$Status) 
```

 Inference - Overall, John is better than Bob

 Day wise Result
```{r}
table(df$Assignee, df$Status, df$Day)
```

 Inference - 
 Day 1 - Bob resolved at least 1 ticket. John has pending ticket. Hence, Bob is better.
 Day 2 - Bob again resolved the ticket. John has pending ticket once again.  Hence, Bob is better.
Day wise, Bob is better.

Simpson's Paradox states that a trend appears in different group of data when seen individually but disappears when these groups are combined.


------------------------------------------------------
 Entropy
------------------------------------------------------

```{r}
library(entropy)
```


Machine 1 entropy
```{r}
entropy.empirical(c(4, 4, 4, 4), unit = "log2")
```


Machine 2 entropy
```{r}
entropy.empirical(c(4, 2, 1, 1), unit = "log2")
```


entropy tends to be more when the probability of all the possible outcomes are close to each other.


------------------------------------------------------
# KL Divergence
------------------------------------------------------
 KL Divergence is the measure of difference between two probability distributions (P.D.).
 KL divergence of P.D. Q with respect to P.D. P gives the measure of the 
information lost by using P to approximate Q.

##### http://ilpce/ce/Course/EXDA/page/169        # definition
##### http://ilpce/ce/Course/EXDA/page/171        # example

```{r}
P <- rep(166, 6)                 
Q <- c(91, 157, 275, 274, 117, 85)
KL.empirical(P, Q, unit = "log2")
```

