---
title: "From forum"
author: "Jeswin"
date: "25 September 2017"
output: html_document
---

```{r}
Data <- structure(list(label = structure(
  c(1L, 1L, 2L, 2L, 3L, 3L, 4L, 
  4L, 5L, 5L, 6L, 6L), .Label = c("A", "B", "C", "D", "E", "F"), class = "factor"), 
  facet = structure(c(1L, 2L, 1L, 2L, 1L, 2L, 1L, 2L, 1L, 2L, 
  1L, 2L), .Label = c("A", "B"), class = "factor"), value = c(0.0108889081049711, 
  0.37984336540103, 0.0232500876998529, 0.777756493305787, 
  0.0552913920022547, 0.920194681268185, 0.0370863009011373, 
  0.114463779143989, 0.00536034172400832, 0.469208759721369, 
  0.0412159096915275, 0.587875489378348), group = c(1, 1, 1, 
  1, 1, 1, 1, 1, 1, 1, 1, 1)), .Names = c("label", "facet", 
  "value", "group"), row.names = c(NA, -12L), class = "data.frame")

head(Data)
```

```{r}
str(Data)
```


I am trying to make a faceted plot in ggplot2 where the y axis shows labels and the x axis should show line graphs with the value for each label in two different measures (which are on different scales).
```{r}
ggplot(Data, aes(x = label, y = value, group = group)) + geom_line() + 
     facet_grid(~ facet, scales = "free") + coord_flip()
```

The problem is that the measures are on different scales and I would prefer the A plot to have x limits from 0 to 0.1 and the B plot to have x limits from 0 to 1. I thought scales = "free" should fix this but it doesn't change the plot.

#### Solu :

Note you have to use __geom_path__, and take care with the ordering of your points because just switching x and y is not the same as __coord_flip__ (which as noted in the other answer isn't supported with facet_wrap).

```{r}
ggplot(Data, aes(y = label, x = value, group=group)) + geom_path() + 
  facet_wrap( ~ facet, scales = "free")
```



























