---
title: "L6-Problem_set_2_variables"
author: "Jeswin"
date: "27 September 2017"
output: html_document
---

```{r}
# Loading libraries
library(ggplot2)
library(tidyr)
suppressMessages(library(dplyr))
suppressMessages(library(gridExtra))
library(hexbin)
```

## 1. Price vs x : create a scatterplot of price vs x.

```{r}
data("diamonds")
summary(diamonds)
```

```{r}
str(diamonds)
```

```{r}
names(diamonds)
```

```{r}
ggplot(data = diamonds, aes(x = x, y = price)) +
  geom_point(alpha = 1/20)
```

#### Notice some outliers and an exponential relationship between price and x

## 2. Find correlation

```{r}
with(diamonds, cor.test(price,x))
```

```{r}
with(diamonds, cor.test(price,y))
```

```{r}
with(diamonds, cor.test(price,z))
```

## 3. Create a simple scatter plot of price vs depth.

```{r}
ggplot(data = diamonds, aes(x = depth, y = price))+
  geom_point(alpha = 1/20) 
  
```


## 5. Adjustments - price vs. depth

Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units.

```{r}
ggplot(data = diamonds, aes(x = depth, y = price))+
  geom_point(alpha = 1/100)+
  scale_x_continuous(breaks = seq(54,70,2), limits = c(54,70))
```

## 7. Correlation - price and depth

```{r}
with(diamonds, cor.test(price,depth))
```

## 8 price vs. carat

Create a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot(data = diamonds, aes(y=price, x=carat))+
  geom_point()+
  ylim(0, quantile(diamonds$price, 0.99))+
  xlim(0, quantile(diamonds$carat, 0.99))
```


## 9. price vs. volume

Create a scatterplot of price vs. volume (x * y * z). Create a new variable for volume in the diamonds data frame.

```{r}
diamonds <- diamonds %>% 
  mutate(volume = x * y * z)
```

```{r}
str(diamonds)
```

```{r}
ggplot(data = diamonds, aes(x = volume, y = price))+
  geom_point()

```

Did you notice some outliers? Some volumes are 0! There's an expensive diamond with a volume near 4000, and a cheaper diamond with a volume near 900.

You can find out how many diamonds have 0 volume by using count(diamonds$volume == 0). The count() function comes with the plyr package.


Note: If you ran the count function from plyr, you need to run this command in R to unload the plyr package.

```
detach("package:plyr", unload=TRUE)
```

The plyr package will conflict with the dplyr package in later exercises.

Depending on your investigation, it may or may not be important for you to understand how outliers, like these, came to be in your data.


```{r}
ggplot(data = diamonds, aes(x = volume, y = price))+
  stat_binhex()+
  scale_fill_gradient(low = 'lightblue', high = 'red')+
  xlim(0,800)
```

```{r}
subd <- subset(diamonds, volume>0 & volume<800 )
with(subd, cor.test(price,volume))
```

## Adjustments - price vs. volume

Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.

```{r}
ggplot(data = subd, aes(y = price, x = volume))+
  geom_point(alpha = 1/20)+
  geom_smooth(method = 'lm', formula = y~poly(x,2) )
```



```{r}
require(mgcv)
library(methods)
```

```{r}
ggplot(data = subd, aes(y = price, x = volume))+
  geom_point(alpha = 1/20)+
  stat_smooth(method = 'gam', formula = y~s(x), size=1 )  # small s
```


## 13 Mean price by clarity

Use the function dplyr package to create a new data frame containing  info on diamonds by clarity.

Name the data frame diamondsByClarity
The data frame should contain the following variables in this order.

(1) mean_price
(2) median_price
(3) min_price
(4) max_price     
(5) n

where n is the number of diamonds in each level of clarity.


```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price), 
            min_price = min(price),
            max_price = max(price),
            n = n())

head(diamondsByClarity)
```



## 14 Bar Charts of Mean Price

We've created summary data frames with the mean price by clarity and color. You can run the code in R to  verify what data is in the variables diamonds_mp_by_clarity  and diamonds_mp_by_color.

Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

diamonds_mp_by_cut <- diamonds %>%
  group_by(cut) %>%
  summarise(mean_price = mean(price))
```

```{r}
p1 <- ggplot(data = diamonds_mp_by_clarity, aes(x = clarity, y = mean_price)) +
  geom_col()

p1
```

```{r}
p2 <- ggplot(data = diamonds_mp_by_color, aes(x = color, y = mean_price)) +
  geom_col()
p2
```
```{r}
p3 <- ggplot(data = diamonds_mp_by_cut, aes(x = cut, y = mean_price))+
  geom_col()

p3
```

```{r}
grid.arrange(p1,p2,p3,ncol=1)
```

Observe : think something odd is going here. These trends seem to go against our intuition.

__Mean price tends to decrease as clarity improves. The same can be said for color.__

***

## 16 [Gapminder Revisited](http://www.gapminder.org/data/)

examine pairs of variable and create 2-5 plots that make use of the techniques from Lesson 4.

```{r}
# Loading libraries
suppressMessages(library(xlsx))
library(rJava)
library(xlsxjars)
```


```{r}
female_bmi <- read.xlsx('Indicator_BMI female ASM.xlsx', sheetName = 'Data', header = TRUE)
male_bmi <- read.xlsx('Indicator_BMI male ASM.xlsx', sheetName = 'Data', header = TRUE)
```

### Selecting countries in which i am interested

```{r}
fm_bmi <- filter(female_bmi, Country %in% c('Bangladesh', 'India', 'Pakistan', 'Russia', 'United Kingdom', 'United States', 'United Arab Emirates'))
m_bmi <- filter(male_bmi, Country %in% c('Bangladesh', 'India', 'Pakistan', 'Russia', 'United Kingdom', 'United States', 'United Arab Emirates') )
#rm(female_bmi,male_bmi)

```

### Selecting only India

```{r}
fm_india <- subset(fm_bmi, Country == 'India')
fm_india_tr <- data.frame(t(fm_india))
rm(fm_india)

m_india <- subset(m_bmi, Country == 'India')
m_india_tr <- data.frame(t(m_india))
rm(m_india)
```


```
# Alternative for turning rownames into a column
# load data table
library(data.table)
# create datable from dataframe
setDT(transposed_data, keep.rownames = TRUE)[]
# rename the first column to 'Year'
names(transposed_data)[[1]] <- "Year"
```
#### [Difference between data.frame() and as.data.frame()](https://stackoverflow.com/questions/21574250/difference-between-as-data-framex-and-data-framex)
#### [How to change a character of each elememt of a column](https://stackoverflow.com/questions/21187603/replace-characters-from-a-column-of-a-data-frame-r)
    Tried to replace 'X1998' to '1998'

## Female BMI india
```{r}
library(data.table)
setDT(fm_india_tr, keep.rownames = TRUE)[]
names(fm_india_tr)[[1]] <- "Years"
fm_india_tr <- fm_india_tr[2:30, ]
fm_india_tr$Years <- gsub('X','',fm_india_tr$Years)   # to remove 'X' from each element of column from 'X2000' to '2000'
names(fm_india_tr)[[2]] <- "Female_BMI"
head(fm_india_tr)
```

```{r}
str(fm_india_tr)
```

[To convert from a factor array to a numeric array or conversion b/w different data types](http://www.cookbook-r.com/Manipulating_data/Converting_between_vector_types/)
The way to get the text values converted to numbers is to first convert it to a character, then a numeric vector.

```{r}
fm_india_tr$Years <- as.numeric(as.character(fm_india_tr$Years))
fm_india_tr$Female_BMI <- as.numeric(as.character(fm_india_tr$Female_BMI))
str(fm_india_tr)
```

## Male BMI India

```{r}
setDT(m_india_tr, keep.rownames = TRUE)[]
names(m_india_tr)[[1]] <- "Years"
m_india_tr <- m_india_tr[2:30, ]
m_india_tr$Years <- gsub('X','',m_india_tr$Years)   # to remove 'X' from each element of column from 'X2000' to '2000'
names(m_india_tr)[[2]] <- "Male_BMI"
head(m_india_tr)
```

```{r}
m_india_tr$Years <- as.numeric(as.character(m_india_tr$Years))
m_india_tr$Male_BMI <- as.numeric(as.character(m_india_tr$Male_BMI))
str(m_india_tr)
```

## Merging the 2 dataframes m_india_tr and fm_india_tr to India_BMI with help of Years column

```{r}
India_BMI <- merge(m_india_tr, fm_india_tr, by = "Years")
head(India_BMI)
```
## Making line graph with multiple lines

```{r}
Indiag_BMI <- gather(India_BMI, key = 'Gender', value = 'BMI', Male_BMI, Female_BMI)
head(Indiag_BMI,10)
```

```{r}
ggplot(data = Indiag_BMI, aes(x = Years, y = BMI, group = Gender, color = Gender))+
  geom_line()+
  geom_point()+
  scale_x_continuous(limits = c(1980,2008), breaks = seq(1980,2008,2))

```

From the figure we observe that in India, women seem to have higher BMI than male. Though that was not the case in earlier 1980s. After 1980s BMI of male population started declining rapidly whereas That of female population in INdia has been on constant rise. After 1991 there has been improvement in the BMI of male population.

### Plotting BMI for males of selected countries 

__m_bmi dataframe__  

```{r}
m_bmi <- m_bmi %>%
  gather(key = 'Years', value = 'BMI', 2:30 )
m_bmi$Years <- gsub('X','',m_bmi$Years)
```

Comparing BMI of males of the selected countries

```{r fig.width = 12, fig.height= 7}
ggplot(data = m_bmi, aes(x = Years, y = BMI))+
  geom_line(aes(color = Country, group = Country))

#ggsave('male_bmi.png')
```

From the graph, we see that US, UAE and UK have the highest male BMI and it is increasing at a steady rate from 1980.
In case of Russia the BMI of males remained almost constant till 2000 and then it started increasing .
COuntries like Pakistan, India and Bangladesh had comparatively lower BMI.

#### Comparing BMI of females of the selected countries

__fm_bmi__

```{r}
fm_bmi <- fm_bmi %>%
  gather(key = 'Years', value = 'BMI', 2:30)
fm_bmi$Years <- gsub('X','',fm_bmi$Years)
```


```{r fig.width = 12, fig.height= 7}
ggplot(data = fm_bmi, aes(x = Years, y = BMI))+
  geom_line(aes(color = Country, group = Country))
#ggsave('female_bmi.png')
```

For female :
From fig, BMI of UAE is higher than other countries. Earlier in 1980s BMI of Russia was higher than that of US but after 1992 BMI of US increased at a ateady rate as compared to that of russia whose BMI decreased comparatively during that time and is increasing steadily after that.


# Additional Graphs

### What is the BMI over time all over the world?

Converting to long format form.

```{r}
male_bmi <- male_bmi %>%
  gather(key = 'Years', value = 'BMI', 2:30)
male_bmi$Years <- gsub('X','',male_bmi$Years)

female_bmi <- female_bmi %>%
  gather(key = 'Years', value = 'BMI', 2:30)
female_bmi$Years <- gsub('X','',female_bmi$Years)

```

### For males
```{r fig.width=12, fig.height=11}
ggplot(data = male_bmi, aes(x = BMI)) +
  geom_histogram(color = 'black', fill = 'white')+
  facet_wrap( ~ Years, ncol = 4)+
  labs(title = 'BMI over the years around the world', x = 'Years', y =  'Count(BMI)' )

```

Over the years, the histograms seem to follow a bimodal distribution.


Also refer [full_join, semi_join, etc ](https://rpubs.com/NateByers/Merging)
Also refer this [blog more important egs](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html)

full_join() includes all observations from x and y dataframes.

```{r}
world_bmi <- full_join(male_bmi, female_bmi)
```

### Global BMI Population between 1980 - 2008.

```{r}
ggplot(data = world_bmi, aes(x = BMI))+
  geom_histogram(color = 'black', fill = 'lightblue')
```




```{r}
ggplot(data = world_bmi, aes(x = BMI))+
  geom_freqpoly(aes(group = Gender, color = Gender))
```

```{r}
ggplot(data = world_bmi, aes(x = Gender, y = BMI, color = Gender))+
  geom_boxplot()+
  scale_y_continuous(limits = c(20,30))
```


Women seem to have a slightly higher BMI overall than males.



```{r fig.width=10, fig.height==8}
ggplot(data = world_bmi, aes(x = Years, y = BMI))+
  geom_line(aes(group = Gender, color = Gender))+
  scale_x_discrete(breaks = seq(1980,2008,1))
```

This one is not much helpful

#### Finished

***





















































