---
title: "NA_in_R"
author: "Jeswin"
date: "14 October 2017"
output: html_document
---

# Missing Data

[Quick R:NA](http://www.statmethods.net/input/missingdata.html)  

[cleaning data in R](https://www.datacamp.com/courses/cleaning-data-in-r)

In R, missing values are represented by the symbol __NA__ (not available). 
Impossible values (e.g., dividing by zero) are represented by the symbol __NaN__ 
(not a number).


## Testing for Missing Values

> is.na(x) # returns TRUE if x is missing

```{r}
y <- c(1,2,3,NA)
is.na(y) # returns a vector (F F F T)
```

## Recoding Values to Missing

```{r}
y[is.na(y)] <- 99
y[1:3]<-NA
y
```

## Excluding Missing Values from Analyses

Arithmetic functions on missing values yield missing values.

```{r}
x <- c(1,2,NA,3)
mean(x)
```

```{r}
mean(x, na.rm=TRUE)  # Removes NA whlie calculating mean
```


```{r}
mydata <- data.frame(
  name = c('Ram', 'Sita', 'Laxman', 'sidekick'),
  Arrows = c(100, NA, 90,2),
  Knife = c(NA, 1 , NA,3),
  God = c(1,1,1,5),
  va = c(1,1,1,66)
)
mydata
```

```{r}
complete.cases(mydata)
```

The function __complete.cases()__ returns a logical vector indicating which 
cases are complete.

```{r}
# list rows of data that have missing values 
mydata[!complete.cases(mydata),]
```

The function __na.omit()__ returns the object with listwise deletion of missing values.

```{r}
# create new dataset without missing data 
newdata <- na.omit(mydata)
newdata
```

## Different blog

[Dealing with Missing Values](http://uc-r.github.io/missing_values)

A common task in data analysis is dealing with missing values. In R, missing 
values are often represented by `NA` or some other value that represents missing 
values (i.e. 99). We can easily work with missing values and in this section 
you will learn how to:

# Test for missing values

To identify missing values use `is.na()` which returns a logical vector with TRUE 
in the element locations that contain missing values represented by NA.`is.na()`
will work on vectors, lists, matrices, and data frames.

```{r}
x <- c(1:4, NA, 6:7, NA)
x
```

```{r}
is.na(x)
```

```{r}
# data frame with missing data
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)
df
```

```{r}
# identify NAs in full data frame
is.na(df)
```

```{r}
# identify NAs in specific data frame column
is.na(df$col4)
```

To identify the location or the number of NAs we can leverage the `which()` 
and `sum()` functions:

```{r}
# identify location of NAs in vector
which(is.na(x))
```

```{r}
# identify count of NAs in data frame
sum(is.na(df))
```

For data frames, a convenient shortcut to compute the total missing values in 
each column is to use `colSums()`:

```{r}
colSums(is.na(df))
```

# Recode missing values

To recode missing values; or recode specific indicators that represent missing 
values, we can use normal subsetting and assignment operations.  

For example, we can recode missing values in vector __x__ with the mean values 
in __x__ by first subsetting the vector to identify NAs and then assign these 
elements a value. 


Similarly, if missing values are represented by another value (i.e. 99) we can 
simply subset the data for the elements that contain that value and then assign 
a desired value to those elements.

```{r}
# recode missing values with the mean
# vector with missing data

x <- c(1:4, NA, 6:7, NA)
x
```

```{r}
x[is.na(x)] <- mean(x, na.rm = TRUE)
round(x,2)
```

```{r}
# data frame that codes missing values as 99
df <- data.frame(col1 = c(1:3, 99), col2 = c(2.5, 4.2, 99, 3.2))
df
```

```{r}
# change 99s to NAs
df[df == 99] <- NA
df
```


If we want to recode missing values in a single data frame variable we can 
subset for the missing value in that specific variable of interest and then 
assign it the replacement value. For example, here we recode the missing value 
in col4 with the mean value of col4.

```{r}
# data frame with missing data
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)
df
```

```{r}
df$col4[is.na(df$col4)] <- mean(df$col4, na.rm = TRUE)
df
```

# Exclude missing values

We can exclude missing values in a couple different ways. First, if we want to 
exclude missing values from mathematical operations use the `na.rm = TRUE` 
argument. If you do not exclude these values most functions will return an NA.

```{r}
# A vector with missing values
x <- c(1:4, NA, 6:7, NA)

# including NA values will produce an NA output
mean(x)
## [1] NA
```

```{r}
# excluding NA values will calculate the mathematical operation for all non-missing values
mean(x, na.rm = TRUE)
```

We may also desire to subset our data to obtain complete observations, those 
observations (rows) in our data that contain no missing data. We can do this a 
few different ways.

```{r}
# data frame with missing values
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)

df
```

First, to find complete cases we can leverage the `complete.cases()` function 
which returns a logical vector identifying rows which are complete cases. So in 
the following case rows 1 and 3 are complete cases. We can use this information 
to subset our data frame which will return the rows which `complete.cases()`
found to be TRUE.

```{r}
complete.cases(df)
```

```{r}
# subset with complete.cases to get complete cases
df[complete.cases(df), ]
```

```{r}
# or subset with `!` operator to get incomplete cases
df[!complete.cases(df), ]
```

An shorthand alternative is to simply use `na.omit()` to omit all rows 
containing missing values.

```{r}
# or use na.omit() to get same as above
na.omit(df)
```












