```{r}
library(ggplot2)
library(MASS)
library(grid)
library(gridExtra)
library(reshape2)
```



## Boxplot

plot them alongside each other
```{r}
op<-par(no.readonly = TRUE)
par(mfrow=c(1,3))
boxplot(trees$Girth)
title(main="Girth of trees",ylab="Tree diameter in inches")
boxplot(trees$Height)
title(main="Height of trees",ylab="Tree height in ft")
boxplot(trees$Volume)
title(main="Volume of trees",ylab="Tree volume in cubic ft")
par(op)
```

plotting them horizontally

```{r}
op<-par(no.readonly = TRUE)
par(mfrow = c(2,1))
boxplot(Cars93$Horsepower, horizontal = T)
title(main="Horsepower of car",ylab="Horsepower")
boxplot(Cars93$MPG.city, horizontal = T)
title(main="MPG car",ylab="MPG")
par(op)
```

```{r}
head(trees)
```

Plotting boxplot of all columns in one graph

```{r}
boxplot(trees)
```

combining different boxplot in one graph

```{r}
p1 <- ggplot() +
  geom_boxplot(data = trees, aes(factor(0), Girth)) +
  scale_x_discrete(breaks = NULL) +
  xlab(" ") +
  ggtitle("Girth of trees")

p2 <- ggplot() +
  geom_boxplot(data = trees, aes(factor(0), Height)) +
  scale_x_discrete(breaks = NULL) +
  xlab(" ") +
  ggtitle("Height of trees")
  
p3 <- ggplot() +
  geom_boxplot(data = trees, aes(factor(0), Volume)) +
  scale_x_discrete(breaks = NULL) +
  xlab(" ") +
  ggtitle("Volume of trees")

grid.arrange(p1,p2,p3,ncol=3)
```

```{r}
# Scatter plot
plot(Cars93$Horsepower, Cars93$MPG.city, type = 'p', pch = "*", cex = 0.9)
```

p -points

l -lines

o - plot points and lines


## scatter plots using the trees data set.
```{r}
plot(trees)
```

Not getting this correctly :

```{r}
attach(Cars93)
Cars93_ordered <- Cars93[order(Horsepower),]
detach(Cars93)
plot(Cars93_ordered$Horsepower, Cars93_ordered$MPG.city, col = "red", lty = 4, ann = F)
lines(Cars93_ordered$Horsepower, Cars93_ordered$MPG.city, col = "blue",lwd = 1, ann = F)
legend("topright", legend = c("City Mileage", "Highway Mileage"), lty = c(4,1), col = c("red", "blue"))
title(main = "Mileage vs Horsepower", xlab = "Horsepower", y = "City Mileage")
```


Trying using ggplot

```{r}
ggplot(data = Cars93_ordered) +
  geom_point(aes(x = Horsepower, y = MPG.city)) +
  geom_line(aes(x = Horsepower, y = MPG.city))
```

***

```{r}
library(ggplot2)
```


```{r}
rm(list = ls())
data(WorldPhones)
df <- WorldPhones  # Is loaded as matrix change to data.frame
head(df)
```

```{r}
class(df)
```

```{r}
df <- data.frame(df)
class(df)
```

```{r}
df$Years <- as.numeric(rownames(df))
df$Years
```

```{r}
df <- subset(df, select = c("Years" ,"N.Amer", "Asia", "Europe"))
colnames(df)
```

As we need only 3 countries so melt the above data frame to use only 3 countries and put it in one column along with values in different column

```{r}
df_melt <- melt(df, id = c("Years"))
head(df_melt)
```

```{r}
colnames(df_melt) <- c("Years", "Countries", "NoOfPhones")
colnames(df_melt)
```


```{r}
ggplot(data = df_melt, aes(x = Years, y = NoOfPhones, group = Countries, colour = Countries)) +
  geom_line()
```


***

```{r}
rm(list = ls())
library(MASS)
attach(Cars93)
df <- Cars93
```


```{r}
head(df)
```

### dot plot
is a representation of distribution of a group of data points plotted on a simple scale. 

```{r}
dotchart(df$Rear.seat.room)
```

```{r}
dotchart(iris$Petal.Length)
title(xlab = "Petal length", main = "Dotchart of petal length")
```

## Bar Plot

```{r}
library(RColorBrewer)
# to use predefined color schemes from this package
```

Let us now find pattern between type of car and its seating capacity. For this purpose, we will use bar chart or bar graph, which is a chart that presents grouped data with rectangular bars with lengths proportional to the values that they represent.

```{r}
table_a <- table(df$Type, df$Passengers)
class(table_a)
table_a
```

```{r}
barplot(table_a, legend = rownames(table_a), col = brewer.pal(6, "Blues"))
title(main = "Number of passengers per car type", xlab = "Number of passenegrs")
```

Color schemes like shown in the figure(continuous shades) should be used when we are representing continuous values.

##### So to represent discrete values using discrete color schemna 

```{r}
barplot(table_a, legend = rownames(table_a), col = brewer.pal(6, "Pastel1"))
title(main = "Number of passengers per car type", xlab = "Number of passenegrs")
```

We observe that with the exception of few, most compact and mid size cars can accommodate 5 people. Here the height signifies the count of different car types in that passenger range.

### Stacked bargraph in ggplot2

 `new_df <- data.frame(table(df$Type, df$Passengers))` when used this data frame is not coming properly so going with table

```{r}
table_a
```

```{r}
new_df <- data.frame(table_a)
head(new_df)
```



```{r}
colnames(new_df) <- c("CarType", "NumberOfPassenegrs", "Count")
head(new_df)
```

```{r}
ggplot(data = new_df, aes(x = NumberOfPassenegrs, y = Count, fill = CarType)) +
  geom_bar(stat = "identity")
```


***

```{r}
rm(list = ls())
library(ggplot2)
library(reshape2)
```

```{r}
data(WorldPhones)
df <- WorldPhones  # Is loaded as matrix change to data.frame
df <- data.frame(df)

df$Years <- as.numeric(rownames(df))
df_melt <- melt(df, id = c("Years"))
colnames(df_melt) <- c("Years", "Countries", "NoOfPhones")

head(df_melt)
```

The graph below is wrong because of `stat = identity` and we had not summed up the values for different countries so need to use table(). 

```{r}
ggplot(data = df_melt, aes(x = Years, y = NoOfPhones, fill = Countries)) +
  geom_bar(stat = "identity") +
  ggtitle("Wrong plot")
```

```{r}
ta <- table(df_melt$Years, df_melt$Countries)
ta
```

```{r}
library(dplyr)
```

```{r}
df2 <- group_by(df_melt, Years, Countries)
df2
```


#### wrong again

```{r}
ggplot(data = df2, aes(x = Years, y = NoOfPhones, fill = Countries)) +
  geom_bar(stat = "identity")
```


#### try using table as in ilpce

```{r}
head(df, 10)
```

```{r}
data(WorldPhones)
df <- WorldPhones  # Is loaded as matrix change to data.frame
df <- data.frame(df)

df$Years <- rownames(df)  # changed the years to factors again
df_melt <- melt(df, id = c("Years"))
colnames(df_melt) <- c("Years", "Countries", "NoOfPhones")

head(df_melt)
```

SO the error was i put the years as numeric but for bargraph it must be as factors

```{r}
ggplot(data = df_melt, aes(x = Years, y = NoOfPhones, fill = Countries)) +
  geom_bar(stat = "identity")
```

##### OR

```{r}
library(RColorBrewer)
barplot(t(WorldPhones),  col = brewer.pal(6, "Pastel1"))
legend("topleft", legend = colnames(WorldPhones), 
       col = brewer.pal(6, "Pastel1"), 
       pch = 15, cex = 0.5)
```

***

### Histogram

```{r}
rm(list = ls())
library(MASS)
data(Cars93)
hist(Cars93$MPG.city, freq = T, breaks = seq(14, 50,2))
```

__freq__ if TRUE, the histogram is a representation of frequencies, else it would plot probability densities.

```{r}
hist(Cars93$MPG.city, freq = F)
```


If we want to draw a probability density plot, then using the density() and lines() we can overlay probability density curve also as seen in Probability Distribution course.

 
```{r}

plot(density(Cars93$MPG.city))
lines(Cars93$MPG.city)
```
***

## Stem and Leaf Chart

```{r}
stem(Cars93$MPG.city)
```


*** 
### BOxplot

```{r}
ggplot() +
  geom_boxplot(data = Cars93, aes(factor(0), Price)) +
  facet_wrap(~Type, scales = "free") +
  scale_x_discrete(breaks = NULL) +
  xlab("") +
  ggtitle("Price of cars according to type")
```

***

## Pie chart

```{r}
ggplot() +
  geom_bar(data = Cars93, aes(x = factor(1), fill = Cylinders), width = 1, position = "fill") +
  xlab("") +
  ylab("") +
  coord_polar(theta = "y") +
  facet_wrap(~Type) +
  scale_fill_brewer(type = "qual", palette = "Pastel1") +
  theme_light() +
  ggtitle("Number of cars in each type according to number of cylinders")
```

***

```{r}
rm(list = ls())
```

```{r}
library(ggplot2)
dow_jones_index <- read.csv("dow_jones_index.data")
head(dow_jones_index)
```

```{r}
dow_jones_index <- subset(dow_jones_index, select = c("stock", "date", "percent_change_price", "percent_change_volume_over_last_wk"),
                          subset = dow_jones_index$stock %in% c("BAC", "AXP"))

dow_jones_index[, "date"] <- as.Date(dow_jones_index[, "date"], "%m/%d/%Y")
head(dow_jones_index)
```

```{r}
a <- ggplot() +
  geom_bar(data = dow_jones_index, aes(x=date, y=percent_change_price, fill = stock), stat = "identity", position = position_dodge())

a + ggtitle("Percentage change in stock price for AXP and BAC") +
  xlab("time") + ylab("Percentage change in stock price")
```

***

## Stock market scenario - Bar Chart and Line Chart

```{r}
colnames(dow_jones_index)
```


a <- ggplot() +
  geom_bar(data = dow_jones_index, aes(x=date, y=percent_change_price, fill = stock), stat = "identity", position = position_dodge()) +
  geom_line(aes(x = date, y = percent_change_volume_over_last_wk, col = stock))



a + ggtitle("Percentage change in stock price for AXP and BAC") +
  xlab("time") + ylab("Percentage change in stock price")

***

```{r}
rm(list = ls())
df <- data.frame(
  group <- c("M", "F", "C"),
  value <- c(25,25, 50)
)

ggplot(df, aes(x = "", y = value, fill = group)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0)
```

***

A.

```{r}
ggplot() + 
  geom_point(data = iris, aes(x = Petal.Length, y = Petal.Width, col=Species, fill = Species), size =3) +
  labs(fill = "Legend") + theme(text = element_text(size=16))
  
```

B. Correct one

```{r}
ggplot() + 
  geom_point(data = iris, aes(x = Petal.Length, y = Petal.Width, col=Species), size =3) +
  labs(col = "Legend") + theme(text = element_text(size=16))
```

C.

```{r}
ggplot() + 
  geom_point(data = iris, aes(x = Petal.Length, y = Petal.Width, col=Species, fill = Species), size =3) +
  labs(col = "Legend") + theme(text = element_text(size=16))
```

D. Correct

```{r}
ggplot() + 
  geom_point(data = iris, aes(x = Petal.Length, y = Petal.Width, col=Species, fill = Species), size =3) +
  labs(col = "Legend", fill = "Legend") + theme(text = element_text(size=16))
```

	
	
	
