---
title: "Creating a new Dataframe from an existing one"
author: "Jeswin"
date: "21 September 2017"
output: html_document
---

## Loading libraries

```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
```


```{r}
head(diamonds)
```



#1: Create data frame with selected columns using column indices

```{r}
# Displays column carat, cut, depth
dfnew1 <- diamonds[,c(1,2,5)]
head(dfnew1)
```



#2: Create data frame with selected columns using column indices with sequences
```{r}
# Displays column carat, cut, color, depth, price, x
dfnew2 <- diamonds[, c(1:3, 5, 7:8)]
head(dfnew2)
```


#3: Create data frame with selected columns using data.frame command

```{r}
# Displays column carat, cut, color
dfnew3 <- data.frame(diamonds$carat, diamonds$cut, diamonds$color)
names(dfnew3) <- c("Carat", "Cut", "Color")
head(dfnew3)
```


#4: Create data frame using selected columns and column names

```{r}
# Displays column carat, depth, price
dfnew4 <- diamonds[,c("carat", "depth", "price")]
head(dfnew4)
```



#5: Create data frame using subset command and column names
```{r}
# Displays column color, carat, price
dfnew5 <- subset(diamonds, select=c("color", "carat", "price"))
head(dfnew5)
```



#6: Create data frame using subset command and column indices

```{r}
# Displays column carat, cut, color, depth
dfnew6 <- subset(diamonds, select=c(1:3, 5))
head(dfnew6)
```
























