---
title: "Lines(ggplot2"
author: "Jeswin"
date: "1 October 2017"
output: html_document
---

[Blog](http://www.cookbook-r.com/Graphs/Lines_(ggplot2)/)

# Lines(ggplot2)
```{r}
# loading libraries 
library(ggplot2)
library(gridExtra)
suppressMessages(library(dplyr))
```


## With one continuous and one categorical axis

```{r}
# Some sample data
dat <- read.table(header = TRUE, text = '
                  cond  result 
                  Control 10
                  treatment 11.5
                  ')
dat
```

## Lines that go all the way across

These use __geom_hline__ because the y-axis is the continuous one, but it is also possible to use geom_vline (with xintercept) if the x-axis is continuous

```{r}
# Basic bar graph
bp <- ggplot(dat, aes(x=cond, y=result)) +
    geom_bar(position=position_dodge(), stat="identity")

# Add a horizontal line
bp2 <- bp + geom_hline(aes(yintercept = 12))

# Make the line red and dashed
bp3 <- bp + geom_hline(aes(yintercept = 12), color = '#990000', linetype = 'dashed')

grid.arrange(bp,bp2,bp3, ncol=3)

```

## Separate lines for each categorical value

To make separate lines for each bar, use __geom_errorbar__. 
The error bars have no height - 'ymin=ymax' . 
It also seems necessary to specify y for some reason, even though it doesn't do anything.

```{r}
dat$hline <- c(9,12)
dat
```

### [Dodging in ggplot2](http://ggplot2.tidyverse.org/reference/position_dodge.html)

Dodging preserves the vertical position of an geom while adjusting the horizontal position.
By default, dodging preserves the total width. You can choose to preserve the width of each element:

```{r}
# Need to re-specify bp, because the data has changed
p1 <- bp <- ggplot(dat, aes(x=cond, y=result)) +
    geom_bar(position=position_dodge(), stat="identity")

# Draw with separate lines for each bar
p2 <- bp + geom_errorbar(aes(ymax=hline, ymin=hline), colour="#AA0000")

# Make the lines narrower
p3 <- bp + geom_errorbar(width=0.5, aes(ymax=hline, ymin=hline), colour="#AA0000")

grid.arrange(p1,p2,p3,ncol=3)
```

Can get the same result, even if we get the hline values from a second data frame. Define data frame with hline

```{r}
dat_hlines <- data.frame(cond=c("control","treatment"), hline=c(9,12))
dat_hlines
```

```{r}
# The bars are from dat, but the lines are from dat_hlines
bp + geom_errorbar(data=dat_hlines, aes(y = NULL, ymax=hline, ymin=hline), colour="#AA0000")
```

## Lines over grouped bars

It is possible to add lines over grouped bars. In this example, there are actually four lines (one for each entry for hline), but it looks like two, because they are drawn on top of each other. I don't think it's possible to avoid this, but it doesn't cause any problems.

```{r}
dat <- read.table(header=TRUE, text='
     cond group result hline
  control     A     10     9
treatment     A   11.5    12
  control     B     12     9
treatment     B     14    12
')
dat
```

```{r}
# Define basic bar plot
bp <- ggplot(dat, aes(x=cond, y=result, fill=group)) +
    geom_bar(position=position_dodge(), stat="identity")


# The error bars get plotted over one another -- there are four but it looks
# like two
bp2 <- bp + geom_errorbar(aes(ymax=hline, ymin=hline), linetype="dashed")

grid.arrange(bp,bp2, ncol=2)

```

# Lines over individual grouped bars
```{r}
# Define basic bar plot
bp <- ggplot(dat, aes(x=cond, y=result, fill=group)) +
    geom_bar(position=position_dodge(), stat="identity")


bp + geom_errorbar(aes(ymax=hline, ymin=hline), linetype="dashed",
                   position=position_dodge())
```





# With two continuous axes

Sample data used here:

```{r}
dat <- read.table(header=TRUE, text='
      cond xval yval
   control 11.5 10.8
   control  9.3 12.9
   control  8.0  9.9
   control 11.5 10.1
   control  8.6  8.3
   control  9.9  9.5
   control  8.8  8.7
   control 11.7 10.1
   control  9.7  9.3
   control  9.8 12.0
 treatment 10.4 10.6
 treatment 12.1  8.6
 treatment 11.2 11.0
 treatment 10.0  8.8
 treatment 12.9  9.5
 treatment  9.1 10.0
 treatment 13.4  9.6
 treatment 11.6  9.8
 treatment 11.5  9.8
 treatment 12.0 10.6
')
```


```{r fig.width=9}
# The basic scatterplot
sp <- ggplot(dat, aes(x=xval, y=yval, colour=cond)) + geom_point()

# Add a horizontal line
hsp <- sp + geom_hline(aes(yintercept=10))

# Add a red dashed vertical line
vsp <- sp + geom_hline(aes(yintercept=10)) +
    geom_vline(aes(xintercept=11.5), colour="#BB0000", linetype="dashed")

grid.arrange(sp, hsp, vsp, ncol=3)
```


## Drawing lines for the mean

It is also possible to compute a mean value for each subset of data, grouped by some variable. The group means would have to be computed and stored in a separate data frame, and the easiest way to do this is to use the dplyr package. Note that the y range of the line is determined by the data.

```{r}
lines <- dat %>%
  group_by(cond) %>%
  summarise(
    x = mean(xval),
    ymin = min(yval),
    ymax = max(yval)
  )

```

```{r}
# Add colored lines for the mean xval of each group
# Add colored lines for the mean xval of each group
sp + geom_hline(aes(yintercept=10)) +
     geom_linerange(aes(x=x, y=NULL, ymin=ymin, ymax=ymax), data=lines)

```

## Using lines with facets

Normally, if you add a line, it will appear in all facets.

```{r}
# Facet, based on cond
spf <- sp + facet_grid(. ~ cond)
spf

```

```{r}
# Draw a horizontal line in all of the facets at the same value
spf + geom_hline(aes(yintercept=10))
```





































































