---
title: "Notes"
author: "Jeswin"
date: "1 October 2017"
output: html_document
---

```{r}
# loading libraries
library(ggplot2)
library(tidyr)
suppressMessages(library(dplyr))
suppressMessages(library(gridExtra))
suppressMessages(library(gcookbook))
```

```{r}
# laoding data
data("mtcars")
data("cabbage_exp")
summary(mtcars)
```
```{r}
str(mtcars)
```

```{r}
str(cabbage_exp)
```

# Mapping Variables to Colors

__to use a variable (column from a data frame) to control the color of geoms.__

Observe here __geom_col()__ was used instead of geom_bar()

```{r}
ggplot(data = cabbage_exp, aes(x=Date, y=Weight))+
  geom_col(aes(fill=Cultivar), color = 'black', position = 'dodge')
```
In the cabbage_exp example, the variable Cultivar is mapped to fill. The Cultivar column in cabbage_exp is a factor, so ggplot2 treats it as a discrete variable.

***

```{r}
ggplot(data = mtcars, aes(x=wt, y = mpg))+
  geom_point(aes(color = cyl))
```

In the mtcars example, cyl is numeric, so it is treated as a continuous variable. Because of this, even though the actual values of cyl include only 4, 6, and 8, the legend has entries for the intermediate values 5 and 7.

To make ggplot() treat cyl as a categorical variable, you can convert it to a factor in the call to ggplot(), or you can modify the
data so that the column is a character vector or factor
```{r}
mtcars$cyl <- factor(mtcars$cyl)  # Converted that column as a facor variable: check using str

ggplot(data = mtcars, aes(x=wt, y = mpg))+
  geom_point(aes(color = cyl))
```

# Using a Different Palette for a Discrete Variable

to use different colors for a discrete mapped variable.

```{r}
data('uspopage') # Age distribution of population in the United States, 1900-2002
summary(uspopage)
```
```{r}
str(uspopage)
```

#### here we'll use the default palette (hue)

__scale_fill_hue() and  scale_colour_hue()  : Colors evenly spaced around the color wheel (same as discrete)__

```{r}
p <- ggplot(data = uspopage, aes(x=Year, y=Thousands, fill = AgeGroup)) + geom_area()
p
```

```
# These three have the same effect (as shown above)
p
p + scale_fill_discrete()
p + scale_fill_hue()
```

Using a ColorBrewer palette

#### scale_fill_brewer() scale_colour_brewer() ColorBrewer palettes

```{r}
p + scale_fill_brewer()
```

- Changing a palette is a modification of the color (or fill) scale: it involves a change in the mapping from numeric or categorical values to aesthetic attributes. 

- There are two types of scales that use colors: __fill__ scales and __color__ scales.

With __scale_fill_hue()__, the colors are taken from around the color wheel in the HCL (hue-chroma-lightness) color space. 
The default lightness value is 65 on a scale from 0-100. This is good for filled areas, but it's a bit light for points and lines.

To make the colors darker for points and lines, as in Figure below, __set the value of l (luminance/lightness)__ :

```{r fig.width=9}
data('heightweight')
#summary(heightweight)

h <- ggplot(data = heightweight, aes(x=ageYear, y=heightIn, color = sex))+
  geom_point()

h2 <- h + scale_fill_hue(l = 45)+
  labs(title = 'Using scale_fill_hue() and lightness l')

grid.arrange(h,h2,ncol=2)
```

The ColorBrewer package provides a number of palettes. You can generate a graphic showing all of them, as shown:

#### scale_fill_brewer doesnt run when loading the below library

```
library(RColorBrewer)
display.brewer.all()
```

ColorBrewer palettes can be selected by name. For example, this will use the Oranges palette :

```{r}
p + scale_fill_brewer(palette="Oranges")
```


You can also use a __palette of greys__. This is useful for print when the output is in black and white. The default is to start at 0.2 and end at 0.8, on a scale from 0 (black) to 1 (white), but you can change the range

```{r fig.width=11, fig.height=7}
g1 <- p + scale_fill_grey()
g2 <- p + scale_fill_grey(start=0.7, end=0)

grid.arrange(g1,g2,ncol=2)
```

```{r}
p + scale_fill_brewer(type = 'qual')
```

So if type = 'qual' not given the function automatically creates a graph with same color but light intensity decreasing with the ordfer of the factor






































































