---
title: "conversion btw wide 2 long"
author: "Jeswin"
date: "29 September 2017"
output: html_document
---

# Reference: [Converting data between wide and long format](http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/)

There are two sets of methods that are explained below:

- gather() and spread() from the tidyr package. This is a newer interface to the reshape2 package.
- melt() and dcast() from the reshape2 package.

```{r}
library(tidyr)
library(reshape2)
suppressMessages(library(dplyr))
```


# Sample data : wide

```{r}
olddata_wide <- read.table(header=TRUE, text='
 subject sex control cond1 cond2
       1   M     7.9  12.3  10.7
       2   F     6.3  10.6  11.1
       3   F     9.5  13.1  13.8
       4   M    11.5  13.4  12.9
')

# Make sure the subject column is a factor
olddata_wide$subject <- factor(olddata_wide$subject)
olddata_wide
```

```{r}
olddata_long <- read.table(header=TRUE, text='
 subject sex condition measurement
       1   M   control         7.9
       1   M     cond1        12.3
       1   M     cond2        10.7
       2   F   control         6.3
       2   F     cond1        10.6
       2   F     cond2        11.1
       3   F   control         9.5
       3   F     cond1        13.1
       3   F     cond2        13.8
       4   M   control        11.5
       4   M     cond1        13.4
       4   M     cond2        12.9
')
# Make sure the subject column is a factor
olddata_long$subject <- factor(olddata_long$subject)
olddata_long
```

# tidyr

## From wide to long : gather()

```{r}
olddata_wide
```

```
 The arguments to gather():
 - data: Data object
 - key: Name of new key column (made from names of data columns)
 - value: Name of new value column
 - ...: Names of source columns that contain values
 - factor_key: Treat the new key column as a factor (instead of character vector)
 - na.rm : If TRUE, will remove rows from output where the value column in NA.
```
```{r}
data_long <- gather(olddata_wide, key = 'condition', value = 'measurement', control:cond2, factor_key=TRUE)
data_long
```

### Trying using dplyr and tidyr

```{r}
data_long_dplr <- olddata_wide %>%
  gather(key = 'condition', value = 'measurement', control:cond2, factor_key=TRUE)
data_long_dplr
```

- In this example, the source columns that are gathered are specified with __control:cond2__. This means to use all the columns, positionally, between control and cond2. 

- Another way of doing it is to name the columns individually, as in:
```
gather(olddata_wide, condition, measurement, control, cond1, cond2)
```
- If you need to use gather() programmatically, you may need to use variables containing column names. To do this, you should use the gather_() function instead, which takes strings instead of bare (unquoted) column names.

```
keycol <- "condition"
valuecol <- "measurement"
gathercols <- c("control", "cond1", "cond2")

gather_(olddata_wide, keycol, valuecol, gathercols)
```

## From long to wide : Use spread()

```
# The arguments to spread():
# - data: Data object
# - key: Name of column containing the new column names
# - value: Name of column containing values
```

```{r}
data_wide_td <- spread(olddata_long, key = condition, value = measurement)
data_wide_td
```


# reshape2



































