---
title: "ggplot2 histogram"
author: "Jeswin"
date: "5 October 2017"
output: html_document
---

```{r}
library(ggplot2)
library(tidyr)
suppressMessages(library(dplyr))
```

__rep(x,times)__ replicates the values in x.

If times consists of a single integer, the result consists of the whole input 
repeated this many times. If times is a vector of the same length as x (after 
replication by each), the result consists of x[1] repeated times[1] times, x[2] 
repeated times[2] times and so on.


```{r}
set.seed(1234)
df <- data.frame(
  sex = factor(rep(c('F','M'), each = 200)),
  weight = round(c(rnorm(200, mean = 55, sd = 5), rnorm(200, mean = 65, sd = 5)))
)
head(df)
```

```{r}
# Basic histogram
ggplot(df, aes(x=weight)) + geom_histogram()
```

```{r}
# Change the width of bins
ggplot(df, aes(x=weight)) + 
  geom_histogram(binwidth=1)
```

```{r}
# Change colors
p<-ggplot(df, aes(x=weight)) + 
  geom_histogram(color="black", fill="white")
p
```

## Add mean line and density plot on the histogram

- The histogram is plotted with density instead of count on y-axis
- Overlay with transparent density plot. The value of alpha controls the level 
of transparency

```{r}
# Add mean vline
p+geom_vline(aes(xintercept = mean(weight)), color = 'blue', linetype = 'dashed',size=1)
```

```{r}
# Histogrm with density plot
ggplot(df, aes(x=weight)) +
  geom_histogram(aes(y = ..density..), color = 'black', fill = 'white')+
  geom_density(alpha=0.2, fill = "#FF6666")

```

## Change histogram plot line types and colors

```{r}
# Change line type
ggplot(df, aes(x=weight))+
  geom_histogram(color="black", fill="lightblue",linetype="dashed")
```

## Change histogram plot colors by groups

### Calculate the mean of each group :

```{r}
mu <- df %>%
  group_by(sex)%>%
  summarise(mean = mean(weight))
mu
```

### Change line colors

Histogram plot line colors can be automatically controlled by the levels of the
variable sex.

```
Note that, you can change the position adjustment to use for overlapping points on the layer. Possible values for the argument 
position are "identity", "stack", "dodge". Default value is "stack".
```
Here tried with the three positionj types.

```{r}
# Change histogram plot line colors by groups
ggplot(df, aes(x=weight, color=sex)) +
  geom_histogram(fill="white", binwidth = 2)
```

```{r}
# Overlaid histograms
ggplot(df, aes(x=weight, color=sex)) +
  geom_histogram(fill="white", alpha=0.5, position="identity", binwidth = 2)
```

```{r}
# Overlaid histograms
ggplot(df, aes(x=weight, color=sex)) +
  geom_histogram(fill="white", alpha=0.5, position="dodge",binwidth = 2)
```

### # Interleaved histograms

```{r}
# Interleaved histograms
ggplot(df, aes(x=weight, color=sex)) +
  geom_histogram(fill="white", position="dodge")+
  theme(legend.position="top")
```

##### Observe how the ggplot and geom_vline use different dataframes

```{r}
# Add mean lines
p<-ggplot(df, aes(x=weight, color=sex)) +
  geom_histogram(fill="white", position="dodge")+
  geom_vline(data=mu, aes(xintercept=mean, color=sex),
             linetype="dashed")+
  theme(legend.position="top")
p
```


It is also possible to change manually histogram plot line colors using the functions :

- __scale_color_manual()__ : to use custom colors
- __scale_color_brewer()__ : to use color palettes from RColorBrewer package
- __scale_color_grey()__ : to use grey color palettes

```{r}
# Use custom color palettes
p+scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

```{r}
# Use brewer color palettes
p+scale_color_brewer(palette="Dark2")
```

```{r}
# Use grey scale
p + scale_color_grey() + theme_classic() +
  theme(legend.position="top")
```

### Change fill colors

Histogram plot fill colors can be automatically controlled by the levels of sex:

```{r}
# Change histogram plot fill colors by groups
ggplot(df, aes(x=weight, fill=sex, color=sex)) +
  geom_histogram(position="identity")
```


```{r}
p<-ggplot(df, aes(x=weight, fill=sex, color=sex)) +
  geom_histogram(position="identity", alpha=0.5)
p
```

```{r}
# Add mean lines
p+geom_vline(data=mu, aes(xintercept=mean, color=sex),
             linetype="dashed", size=2)
```

Some plots remaining : Refer : ***sthda.com ggplot2 essentials histogram***



















































