---
title: "dplyr packagetutorial"
author: "Jeswin"
date: "22 September 2017"
output: html_document
---

Learnt from [Youtube Datascience video on dplyr](https://www.youtube.com/watch?v=jWjqLW-u3hc)
Markdown link [Notes](http://rpubs.com/justmarkham/dplyr-tutorial)

## Why do I use dplyr?
- Great for data exploration and transformation
- Intuitive to write and easy to read, especially when using the "chaining" syntax (covered below)
- Fast on data frames

## dplyr functionality
* Five basic verbs: 'filter', 'select', 'arrange', 'mutate', 'summarise' (plus 'group_by')
* Can work with data stored in databases and data tables
* Joins: inner join, left join, semi-join, anti-join (not covered below)
* Window functions for calculating ranking, offsets, and more
* Better than plyr if you're only working with data frames (though it doesn't yet duplicate all of the plyr functionality)

## Loading dplyr and an example dataset

* dplyr will mask a few base functions
* hflights is flights departing from two Houston airports in 2011

```{r}
# Load packages

suppressMessages(library(dplyr))
library(hflights)

# explore data

data(hflights)
head(hflights)
```

* 'tbl_df' creates a "local data frame"
* Local data frame is simply a wrapper for a data frame that prints nicely. Otherwise it is same as dataframe.
* It will print only that much which comes on the screen(console) i.e., adjusts what is to be printed.

```{r}
# Convert to local dataframe
flights <- tbl_df(hflights)

# Printing 
flights
```

```{r}
# you can specify that you want to see more rows
print(flights, n=20)

```

```{r}
# convert to a normal data frame to see all of the columns
data.frame(head(flights))
```

## filter: Keep rows matching criteria

- Base R approach to filtering forces you to repeat the data frame's name
- dplyr approach is simpler to write and read
- Command structure (for all dplyr verbs):
     - first argument is a data frame
     - return value is a data frame
     - nothing is modified in place
- Note: dplyr generally does not preserve row names

```{r}
# base R approach to view all flights on January 1

flights[flights$Month==1 & flights$DayofMonth==1, ]
```

### dplyr approach

```{r}
# dplyr approach
# note: you can use comma or ampersand to represent AND condition
# As earlier nothing due to this command is modified inplace, We get to see only a view

filter(flights, Month==1 & DayofMonth == 1)

```

```{r}
# use pipe for OR condition
filter(flights, UniqueCarrier == "AA" | UniqueCarrier == "UA")
```

For the same OR operation __you can also use %in% operator.__

```{r}
# you can also use %in% operator
filter(flights, UniqueCarrier %in% c("AA", "UA"))
```

# select: Pick columns by name
- Base R approach is awkward to type and to read
- dplyr approach uses similar syntax to 'filter'
- Like a SELECT in SQL

```{r}
# base R approach to select DepTime, ArrTime, and FlightNum columns
flights[ , c("DepTime", "ArrTime", "FlightNum")]
```

```{r}
# dplyr approach
select(flights, DepTime, ArrTime, FlightNum)
```

```{r}
names(flights)
```


```{r}
# use colon to select multiple contiguous columns, and use `contains` to match columns by name
# note: `starts_with`, `ends_with`, and `matches` (for regular expressions) can also be used to match columns by name
select(flights, Year:DayofMonth, contains("Taxi"), contains("Delay"))
```

Year:DayofMonth --> It means to select all colimns from 'Year' to 'DayofMonth'.

contains("Taxi") --> Selects all columns having name "Taxi"

contains("Delay") --> Selects all columns having name "Delay"

# "Chaining" or "Pipelining"

- Usual way to perform multiple operations in one line is by nesting
- Can write commands in a natural order by using the '%>%' infix operator (which can be pronounced as "then")

```{r}
# nesting method to select UniqueCarrier and DepDelay columns and filter for delays over 60 minutes
filter(select(flights, UniqueCarrier, DepDelay), DepDelay > 60)
```

```{r}
# Chaining methods
flights %>%
  select(UniqueCarrier, DepDelay) %>%
  filter(DepDelay > 60)

```

- Chaining increases readability significantly when there are many commands
- Operator is automatically imported from the [magrittr](https://github.com/tidyverse/magrittr) package
- Can be used to replace nesting in R commands outside of dplyr

```{r}
# create two vectors and calculate Euclidian distance between them
x1 <- 1:5; x2 <- 2:6
sqrt(sum((x1-x2)^2))
```

Now doing the same code using the '%>%' operator

```{r}
# chaining method
(x1 - x2)^2 %>% sum() %>% sqrt()

```


# arrange: Reorder rows

```{r}
 # base R approach to select UniqueCarrier and DepDelay columns and sort by DepDelay

flights[order(flights$DepDelay), c("UniqueCarrier","DepDelay")]
```

```{r}
# dplyr approach

flights %>%
  select(UniqueCarrier, DepDelay) %>%
  arrange(DepDelay)   # By default arrange does in ascending order

```

```{r}
# dplyr approach
# use `desc` for descending

flights %>%
  select(UniqueCarrier, DepDelay) %>%
  arrange(desc(DepDelay))
```

# mutate: Add new variables

- Create new variables that are functions of existing variables

```
# base R approach to create a new variable Speed (in mph)

flights$Speed < flights$Distance / flights$AirTime*60
flights[, c("Distance", "AirTime", "Speed")]

```

```{r}
# dplyr approach (prints the new variable but does not store it)

flights %>%
  select(Distance, AirTime) %>%
  mutate(Speed = Distance/AirTime*60)

```

```{r}
# store the new variable
# Observe we need not select the variables to use it in mutate function

flights <- flights %>% mutate(Speed = Distance/AirTime*60)
```

# summarise: Reduce variables to values

- Primarily useful with data that has been grouped by one or more variables
- 'group_by' creates the groups that will be operated on
- 'summarise' uses the provided aggregation function to summarise each group

```
# base R approaches to calculate the average arrival delay to each destination
head(with(flights, tapply(ArrDelay, Dest, mean, na.rm=TRUE)))
head(aggregate(ArrDelay ~ Dest, flights, mean))
```

```{r}
# dplyr approach: create a table grouped by Dest, and then summarise each group by taking the mean of #ArrDelay
flights %>%
  group_by(Dest) %>%
  summarise(avg_delay = mean(ArrDelay, na.rm = TRUE))

```

- 'summarise_each' allows you to apply the same summary function to multiple columns at once
- Note: 'mutate_each' is also available
- 'summarise_each(funs(mean), Cancelled, Diverted)' In the code below if we ned to find mean of many more columns instead of writimg each column names we can use functions such as `starts_with`, `ends_with`, and `matches`,etc. 

```{r}
# for each carrier, calculate the percentage of flights cancelled or diverted
# IMP : In qtn if 'for each' is den it means to use 'group_by'
# summarise_each kas keyword funs() : inside which we need to specify the function we need to apply to each column

# As 'Cancelled' and 'Diverted' are columns having values of 0's and 1's so mean of a column will give a %ge of that column

flights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(mean), Cancelled, Diverted)
```

  
```{r}
# for each carrier, calculate the minimum and maximum arrival and departure delays
flights %>%
  group_by(UniqueCarrier) %>%
  summarise_each(funs(min(. , na.rm = TRUE), max(. , na.rm=TRUE)), matches("Delay"))
```

### These functions are found exclusively 
- Helper function n() counts the number of rows in a group
-  Helper function n_distinct(vector) counts the number of unique items in that vector

```{r}
# for each day of the year, count the total number of flights and sort in descending order
flights %>%
  group_by(Month, DayofMonth) %>%
  summarise(flight_count = n()) %>%
  arrange(desc(flight_count))

```

```{r}
# rewrite more simply with the `tally` function
# tally() is equivalent to summarise() + arrange()

flights %>%
  group_by(Month, DayofMonth) %>%
  tally(sort = TRUE)

```


Difference between tally and summarise is that using summarise we can name the count column but in tally it is named as n.

```{r}
# for each destination, count the total number of flights and the number of distinct planes that flew there
flights %>%
  group_by(Dest) %>%
  summarise(flight_count = n(), plane_count = n_distinct(TailNum))


```

* Grouping can sometimes be useful without summarising
```{r}
# for each destination, show the number of cancelled and not cancelled flights

flights %>%
  group_by(Dest) %>%
  select(Cancelled) %>%
  table() %>%
  head()
```

# Window Functions

* Aggregation function: (like mean) Takes n inputs -> Returns 1 output
* [Window function](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html) :  Takes n inputs -> Returns n output
* Window function takes n inputs and returns n values
Includes __ranking and ordering functions (like 'min_rank')__, 
__offset functions ('lead' and 'lag')__, and __cumulative aggregates (like 'cummean')__.
And __Rolling aggregates.__

```{r}
# for each carrier, calculate which two days of the year they had their longest departure delays
# note: smallest (not largest) value is ranked as 1, so you have to use `desc` to rank by largest value

flights %>%
  group_by(UniqueCarrier) %>%
  select(Month, DayofMonth, DepDelay) %>%
  filter(min_rank(desc(DepDelay)) <= 2) %>%
  arrange(UniqueCarrier, desc(DepDelay))
  
```

```{r}
# rewrite more simply with the `top_n` function
# There is a problem in dplyr0.2 dnt know how it selects DepDelay

flights %>%
  group_by(UniqueCarrier) %>%
  select(Month, DayofMonth, DepDelay) %>%
  top_n(2)  %>% 
  arrange(UniqueCarrier, desc(DepDelay))

```


```{r}
# for each month, calculate the number of flights and the change from the previous month
# lag checks the previous months value here
# leads checks the next value

flights %>%
  group_by(Month) %>%
  summarise(flight_count = n())  %>%
  mutate(change = flight_count - lag(flight_count))

```

```{r}
# rewrite more simply with the `tally` function
flights %>%
  group_by(Month) %>%
  tally() %>%
  mutate(change = n - lag(n))   # Observe used n inplace of flight_count

```

# Other Useful Convenience Functions

```{r}
# randomly sample a fixed number of rows, without replacement
flights %>% sample_n(5)
```

```{r}
# randomly sample a fraction of rows, with replacement
flights %>% sample_frac(0.25, replace = TRUE)
```

```{r}
# base R approach to view the structure of an object
str(flights)
```

```{r}
# dplyr approach: better formatting, and adapts to your screen width
glimpse(flights)

```

# Connecting to Databases
- dplyr can connect to a database as if the data was loaded into a data frame
- Use the same syntax for local data frames and databases
- Only generates SELECT statements i.e., cant make any changes
- Currently supports SQLite, PostgreSQL/Redshift, MySQL/MariaDB, BigQuery, MonetDB
- Example below is based upon an SQLite database containing the hflights data
     - Instructions for creating this database are in the databases vignette


```
# connect to an SQLite database containing the hflights data
#my_db <- src_sqlite("my_db.sqlite3")
```

 __problem with using databases , check later__











































