---
title: "dplyr2_tutorial"
author: "Jeswin"
date: "2 October 2017"
output: html_document
---

```{r}
# load packages
suppressMessages(library(dplyr))
library(nycflights13)
```

```{r}
# load the flights dataset from nycflights13
data("flights")
head(flights)
```

```{r}
str(flights)
```

## Choosing columns: select, rename

```{r}
# using select() to pick columns...

flights %>%
  select(carrier,flight)
  
```
```{r}
# To select all columns except the month, day, hour, year, minute and time_hour column

flights %>% select(-month,-day,-hour,-year,-minute,-time_hour)
```

##### Colon operator : can be used to select a range of columns

```{r}
# hide a range of columns
flights %>% select(-(dep_time:time_hour))
```

```{r}
# hide any column with a matching name

flights %>% select(-contains('time'))
```

```{r}
# select() can be used to rename columns, though all columns not mentioned are dropped
# Here renaming tailnum column to tail

flights %>% select(tail = tailnum)
```

```{r}
# rename() does the same thing, except all columns not mentioned are kept
flights %>% rename(YEAR = year)
```

## Choosing rows: filter, between, slice, sample_n, top_n, distinct

```{r}
# filter() supports the use of multiple conditions

flights %>%  filter(dep_time >= 600, dep_time <= 605)
```

```{r}
# between() is a concise alternative for determing if numeric values fall in a range
# Does the same thing as above
flights %>% filter(between(dep_time, 600, 605))
```

```
# side note: is.na() can also be useful when filtering for NA

flights %>% filter(!is.na(dep_time))
```

```{r}
# slice() filters rows by position
flights %>% slice(1000:1005)
```

```{r}
# keep the first three rows within each group
flights %>% 
  group_by(month, day) %>% 
  slice(1:3)
```

```{r}
# sample three rows from each group : sample_n()
flights %>% 
  group_by(month, day) %>% 
  sample_n(3)
```

```{r}
# keep three rows from each group with the top dep_delay
# also sort by dep_delay within each group
flights %>% 
  group_by(month, day) %>% 
  top_n(3, dep_delay) %>%
  arrange(desc(dep_delay))
```


```{r}
# unique rows can be identified using unique() from base R
flights %>% select(origin, dest) %>% unique()
```

```
# dplyr provides an alternative that is more "efficient"
flights %>% select(origin, dest) %>% distinct()

# side note: when chaining, you don't have to include the parentheses if there are no arguments
flights %>% select(origin, dest) %>% distinct
```

## Adding new variables: mutate, transmute, add_rownames

```{r}
# mutate() creates a new variable (and keeps all existing variables)
flights %>% mutate(speed = distance/air_time*60)
```

```{r}
# transmute() only keeps the new variables
flights %>% transmute(speed = distance/air_time*60)
```

```{r}
data('mtcars')
mtcars %>% head()
```

Observe here mtcars uses __car names__ as __row names__ to store the details about the cars. In tidy data principles it is not recommended. Instead the row names must be used as an exclusive variable

So for such cases use the function __add_rownames('new_names')__

```{r}
mtcars %>%
  add_rownames('Models') %>%
  head()
```

```{r}
# side note: dplyr no longer prints row names (ever) for local data frames
mtcars %>% tbl_df()
```

# Grouping and counting: summarise, tally, count, group_size, n_groups, ungroup

```{r}
# summarise() can be used to count the number of rows in each group
flights %>% group_by(month) %>% summarise(cnt = n())
```


```{r}
# tally() and count() can do this more concisely
flights %>% group_by(month) %>% tally()
```

#### count() function does the grouping for us

```{r}
flights %>% count(month)  
```

```{r}
# you can sort by the count
flights %>%
  group_by(month) %>%
  summarise(cnt = n()) %>%
  arrange(desc(cnt))
```

#### tally() and count() have a sort parameter for this purpose

```{r}
flights %>% 
  group_by(month) %>%
  tally(sort = TRUE)
```

```{r}
flights %>% count(month, sort=TRUE)
```


```{r}
# you can sum over a specific variable instead of simply counting rows
flights %>%
  group_by(month) %>%
  summarise(Distance = sum(distance))
```

##### To get the same result as above tally() and count() have a wt parameter
```
flights %>% group_by(month) %>% tally(wt = distance)
flights %>% count(month, wt = distance)
```

```{r}
# group_size() returns the counts as a vector
flights %>% 
  group_by(month) %>%
  group_size(
    
  )

```

```{r}
# n_groups() simply reports the number of groups
flights %>% 
  group_by(month) %>% 
  n_groups()
```

```{r}
# group by two variables, summarise, arrange (output is possibly confusing)

flights %>%
  group_by(month, day) %>%
  summarise(cnt = n()) %>%
  arrange(desc(cnt)) %>%
  print(n=40)  # print 40 rows
```

# Creating data frames: data_frame

'data_frame()' is a better way than 'data.frame()' for creating data frames. 

Benefits of 'data_frame()':

- You can use previously defined columns to compute new columns.
- It never coerces column types.
- It never munges column names.
- It never adds row names.
- It only recycles length 1 input.
- It returns a local data frame (a tbl_df).

```{r}
# data_frame() example
# observe how column b has been created using variable a (this cant be done using data.frame)
data_frame(a = 1:6, b = a*2, c = 'string', 'd+e' = 1) %>%
  glimpse()

```

```{r}
# data.frame() example
data.frame(a = 1:6, c = 'string', 'd+e' = 1) %>% 
  glimpse()
```

# Joining (merging) tables: left_join, right_join, inner_join, full_join, semi_join, anti_join

```{r}
# Create 2 simple dataframes

a <- data_frame(color = c('green','yellow','red'), num = 1:3)
a
```

```{r}
(b <- data_frame(color = c('green','yellow','red'), size = c('S','M','L')))
```

```{r}
# only include observations found in both "a" and "b" (automatically joins on variables that appear in both tables)
inner_join(a, b)
```

```{r}
# include observations found in either "a" or "b"
full_join(a, b)
```

```{r}
# include all observations found in "a"
left_join(a, b)
```

```{r}
# include all observations found in "b"
right_join(a, b)
```

```{r}
# filter "a" to only show observations that match "b"
semi_join(a, b)
```

```{r}
# filter "a" to only show observations that don't match "b"
(anti_join(a, b))
```

Not showing the red row


```{r}
# sometimes matching variables don't have identical names
b <- b %>% rename(col = color)

# specify that the join should occur by matching "color" in "a" with "col" in "b"
inner_join(a, b, by=c("color" = "col"))
```


# Notes on printing 
Refer videos
















































